Faithful Variable Screening for\\ High-Dimensional Convex Regression

### 2014-10-01 tex/scam/additive_faithful.tex
new linear example

@@

\begin{example} 
\label{examp:boundaryflat_example}
 Consider a mixture $\lambda U[-(b+\epsilon), b+\epsilon]^2 + (1-\lambda) N_b(0, \Sigma)$ where $N_b(0,\Sigma)$ is the density of a \emph{truncated} bivariate Gaussian bounded in $[-b, b]^2$. The uniform distribution is supported over a slightly larger square to satisfy the boundary flatness conditions.

@@

\begin{lemma}
\label{cor:faithfulness_extension}
Suppose $p(x)$ is a positive density on $C=[0,1]^p$ satisfying the boundary flatness condition.
Let $f(\mathbf{x})$ be a convex twice differentiable function on $C$. Let $\phi(\mathbf{x}_{-k})$ be any function that does not depend on $x_k$. Then, we have that the unconstrained univariate function
 \begin{align}
h^*_k = \arg\min_{f_k} \mathbb{E} \Bigl[\bigl( f(X) 
           - \phi(X_{-k}) - h_k(X_k) \bigr)^2\Bigr]
\end{align}
is given by $h^*_k(x_k) = \mathbb{E}\bigl[ f(X) - \phi(X_{-k}) \,|\, x_k\bigr]$,
and $ h^*_k = 0$ implies that $\partial_{x_k} f(x) = 0$.
\end{lemma}

>>>>

\begin{remark}
In Theorem~\ref{thm:convex_faithful}, we do not assume a parametric form for the additive components. The additive approximations may not be faithful if we use parametric components. For example, suppose we approximate a convex function $f(X)$ by a linear form $X \beta$. The optimal linear function in the population setting is $\beta^* = \Sigma^{-1} \trm{Cov}(X, f(X))$. Suppose the $X$'s are independent and follow a symmetric distribution and suppose $f(\mathbf{x}) = x_1^2 - \E[X_1^2]$, then $\beta^*_1 = \E[ X_1 f(X)] = \E[ X_1^3 - X_1 \E[X_1^2]] = 0$.
\end{remark}

@@

\begin{example} 
\label{examp:boundaryflat_example}
 Consider a mixture $\lambda U[-(b+\epsilon), b+\epsilon]^2 + (1-\lambda) N_b(0, \Sigma)$ where $N_b(0,\Sigma)$ is the density of a \emph{truncated} bivariate Gaussian bounded in $[-b, b]^2$ and $U[-(b+\epsilon), b+\epsilon]^2$ is the uniform distribution over a square. The uniform distribution is supported over a slightly larger square to satisfy the boundary flatness conditions.

@@

\begin{lemma}
\label{cor:faithfulness_extension}
Suppose $p(x)$ is a positive density on $C=[0,1]^p$ satisfying the boundary flatness condition.
Let $f(\mathbf{x})$ be a convex twice differentiable function on $C$. Let $\phi(\mathbf{x}_{-k})$ be any function that does not depend on $x_k$. Then, we have that the unconstrained univariate function
 \begin{align}
h^*_k = \arg\min_{f_k} \mathbb{E} \Bigl[\bigl( f(X) 
           - \phi(X_{-k}) - h_k(X_k) \bigr)^2\Bigr]
\end{align}
is given by $h^*_k(x_k) = \mathbb{E}\bigl[ f(X) - \phi(X_{-k}) \,|\, x_k\bigr]$,
and $ h^*_k = 0$ implies that $\partial_{x_k} f(\mathbf{x}) = 0$.
\end{lemma}

### 2014-10-01 tex/scam/sparsistency.tex
minor edits
new linear example

\begin{definition}
\label{def:restricted_regression}
We define the \emph{restricted regression} problem 
\[
\min_{d_k} \frac{1}{n} \Big\| Y - \sum_{k \in S} \bar{\Delta}_k d_k \Big\|_2^2 + 
   \lambda_n \sum_{k \in S} \| \bar{\Delta}_k d_k \|_\infty \quad \trm{such that} \, d_{k,1}, \ldots, d_{k,n-1} \geq 0
\]
where we restrict the indices $k$ in
optimization \eqref{opt:alternate_opt} to lie in the set $S$ of true
relevant variables.
\end{definition}

@@

This result holds regardless of whether or not we impose the boundedness conditions in optimization~\eqref{opt:alternate_opt} and~\eqref{opt:alternate_opt_concave}.
The full proof of Theorem~\ref{thm:deterministic} is in Section~\ref{sec:deterministic_proof} of the Appendix.

@@

Let $\mathcal{C}^1$ denote the set of univariate convex functions
supported on $[-b,b]$, 
and let  $\mathcal{C}_1^{p}$ denote the set of convex additive functions
$\mathcal{C}_1^p \equiv \{ f \,:\, f = \sum_{k=1}^p f_k, \,
   f_k \in \mathcal{C}^1 \} $.  
Let $f^*(x) = \sum_{k=1}^p f^*_k(x_k)$ be the population risk
minimizer in $\mathcal{C}_1^p$, 
\begin{equation}
f^* \equiv \arg\min_{f \in \mathcal{C}_1^p} \E\big(f_0(X) - f^*(X)
\big)^2.
\end{equation}
Similarly, we define $\mh \mathcal{C}^1$ as the set of univariate concave functions supported on $[-b, b]$ and define
\begin{equation}
g^*_k = \arg\min_{g_k \in \mh \mathcal{C}^1} \E \big( f_0(X) - f^*(X)
- g_k(X_k) \big)^2.
\end{equation}
We let $S = \{ k = 1,\ldots,p \,:\, f^*_k \neq 0 \trm{ or } g^*_k \neq 0\}$. By additive faithfulness (Theorem~\ref{thm:acdc_faithful}), it must be that $S_0 \subset S$. 
Each of our theorems will use a subset of the following assumptions:
\begin{packed_enum}
\item[A1:] $X_S, X_{S^c}$ are independent. 
\item[A2:] $f_0$ is convex and twice-differentiable. 
\item[A3:] $\|f_0\|_\infty \leq sB$ and $\| f^*_k \|_\infty \leq B$ for all $k$.
\item[A4:] $W$ is mean-zero sub-Gaussian, independent of $X$, with scale $\sigma$; i.e., for all $t \in \R$, $\E e^{t \epsilon} \leq e^{\sigma^2 t^2 / 2}$.
\end{packed_enum}
By assumption A1, $f^*_k$ is must be zero for $k\notin S$.
We define $\alpha_{+}, \alpha_{-}$ as a measure of the signal strength of the weakest variable:
\begin{align*}
\alpha_{+} &= \inf_{f \in \mathcal{C}_1^p \,:\, \exists k ,\, f^*_k \neq 0 \,\wedge\, f_k = 0} 
       \Big\{ \mathbb{E} \big( f_0(X) - f(X) \big)^2 - 
        \mathbb{E} \big( f_0(X) - f^*(X) \big)^2  \Big\}\\
\alpha_{-} &=   \min_{k \in S \,:\, g^*_k \neq 0}
      \Big\{ \mathbb{E} \big( f_0(X) - f^*(X) \big)^2 - 
    \mathbb{E} \big( f_0(X) - f^*(X) - g^*_k(X_k) \big)^2 \Big\}
\end{align*}
These quantities play the role of the absolute value of the smallest nonzero
coefficient in the true linear model in lasso theory.  Intuitively, if
$\alpha_{+}$ is small, then it is easier to make a false omission in the
additive convex stage of the procedure. If $\alpha_{-}$ is small, then
it is easier to make a false omission in the decoupled concave stage
of the procedure.

\begin{remark}
  We make strong assumptions on the covariates in A1 in order to make
  very weak assumptions on the true regression function $f_0$ in
  A2. In particular, we do not assume that $f_0$ is additive. In
  important direction for future work is to weaken this assumption.
  Our simulation experiments indicate that the procedure can be
  effective even when the relevant and irrelevant variables are correlated.
  %Strong assumptions on the covariates are not uncommon in nonparametric
  %variable selection analysis~\cite{lafferty2008rodeo}. 
  %[TODO: refer to correlated design experiment].
\end{remark}

@@

Combining Theorems~\ref{thm:false_positive} and
~\ref{thm:false_negative} 
we obtain the following result.
\begin{corollary}
  Suppose the assumptions of Theorem~\ref{thm:false_positive} and
  Theorem~\ref{thm:false_negative} hold.  
Then with probability at least $1-\frac{C}{n}$
\begin{align}
\hat{f}_k \neq 0 \trm{ or }\; \hat{g}_k \neq 0 &\trm{ for all } k \in S\\
\hat{f}_k = 0 \trm{ and }\; \hat{g}_k = 0 & \trm{ for all } k \notin S
\end{align}
where $c,C$ are constants.

>>>>

\begin{definition}
\label{def:restricted_regression}
We define the \emph{restricted regression} problem 
\[
\min_{d_k} \frac{1}{n} \Big\| Y - \sum_{k \in S} \bar{\Delta}_k d_k \Big\|_2^2 + 
   \lambda_n \sum_{k \in S} \| \bar{\Delta}_k d_k \|_\infty \quad \trm{such that} \, d_{k,1}, \ldots, d_{k,n-1} \geq 0
\]
where we restrict the indices $k$ in
optimization \eqref{opt:alternate_opt} to lie in some set $S$ which contains the true
relevant variables.
\end{definition}

@@

This result holds regardless of whether or not we impose the boundedness conditions in optimization~\eqref{opt:alternate_opt} and~\eqref{opt:alternate_opt_concave}.
The full proof of Theorem~\ref{thm:deterministic} is in Section~\ref{sec:deterministic_proof} of the Appendix. We allow $S$ in Theorem~\ref{thm:deterministic} to be any set containing the relevant variables; in Lasso analysis, $S$ is taken to be the set of relevant variables; we will take $S$ to be the set of variables chosen by the additive convex and decoupled concave procedure in the population setting, which is guaranteed to contain the relevant variables because of additive faithfulness.

@@

Let $\mathcal{C}^1$ denote the set of univariate convex functions
supported on $[-b,b]$, 
and let  $\mathcal{C}_1^{p}$ denote the set of convex additive functions
$\mathcal{C}_1^p \equiv \{ f \,:\, f = \sum_{k=1}^p f_k, \,
   f_k \in \mathcal{C}^1 \} $.  
Let $f^*(\mathbf{x}) = \sum_{k=1}^p f^*_k(x_k)$ be the population risk
minimizer in $\mathcal{C}_1^p$, 
\begin{equation}
f^* \equiv \arg\min_{f \in \mathcal{C}_1^p} \E\big(f_0(X) - f^*(X)
\big)^2.
\end{equation}
Similarly, we define $\mh \mathcal{C}^1$ as the set of univariate concave functions supported on $[-b, b]$ and define
\begin{equation}
g^*_k = \arg\min_{g_k \in \mh \mathcal{C}^1} \E \big( f_0(X) - f^*(X)
- g_k(X_k) \big)^2.
\end{equation}
We let $S = \{ k = 1,\ldots,p \,:\, f^*_k \neq 0 \trm{ or } g^*_k \neq 0\}$. By additive faithfulness (Theorem~\ref{thm:acdc_faithful}), it must be that $S_0 \subset S$. 
Each of our theorems will use a subset of the following assumptions:
\begin{packed_enum}
\item[A1:] $X_S, X_{S^c}$ are independent. 
\item[A2:] $f_0$ is convex and twice-differentiable. 
\item[A3:] $\|f_0\|_\infty \leq sB$ and $\| f^*_k \|_\infty \leq B$ for all $k$.
\item[A4:] $W$ is mean-zero sub-Gaussian, independent of $X$, with scale $\sigma$; i.e., for all $t \in \R$, $\E e^{t \epsilon} \leq e^{\sigma^2 t^2 / 2}$.
\end{packed_enum}
By assumption A1, $f^*_k$ is must be zero for $k\notin S$.
We define $\alpha_{+}, \alpha_{-}$ as a measure of the signal strength of the weakest variable:
\begin{align*}
\alpha_{+} &= \inf_{f \in \mathcal{C}_1^p \,:\, \exists k ,\, f^*_k \neq 0 \,\wedge\, f_k = 0} 
       \Big\{ \mathbb{E} \big( f_0(X) - f(X) \big)^2 - 
        \mathbb{E} \big( f_0(X) - f^*(X) \big)^2  \Big\}\\
\alpha_{-} &=   \min_{k \in S \,:\, g^*_k \neq 0}
      \Big\{ \mathbb{E} \big( f_0(X) - f^*(X) \big)^2 - 
    \mathbb{E} \big( f_0(X) - f^*(X) - g^*_k(X_k) \big)^2 \Big\}
\end{align*}
These quantities play the role of the absolute value of the smallest nonzero
coefficient in the true linear model in lasso theory.  Intuitively, if
$\alpha_{+}$ is small, then it is easier to make a false omission in the
additive convex stage of the procedure. If $\alpha_{-}$ is small, then
it is easier to make a false omission in the decoupled concave stage
of the procedure.

\begin{remark}
  We make strong assumptions on the covariates in A1 in order to make
  weak assumptions on the true regression function $f_0$ in
  A2. In particular, we do not assume that $f_0$ is additive. In
  important direction for future work is to weaken assumption A1.
  Our simulation experiments indicate that the procedure can be
  effective even when the relevant and irrelevant variables are correlated.
  %Strong assumptions on the covariates are not uncommon in nonparametric
  %variable selection analysis~\cite{lafferty2008rodeo}. 
  %[TODO: refer to correlated design experiment].
\end{remark}

@@

Combining Theorems~\ref{thm:false_positive} and
~\ref{thm:false_negative} 
we obtain the following result.
\begin{corollary}
  Suppose the assumptions of Theorem~\ref{thm:false_positive} and
  Theorem~\ref{thm:false_negative} hold.  
Then with probability at least $1-\frac{C}{n}$
\begin{align}
\hat{f}_k \neq 0 \trm{ or }\; \hat{g}_k \neq 0 &\trm{ for all } k \in S\\
\hat{f}_k = 0 \trm{ and }\; \hat{g}_k = 0 & \trm{ for all } k \notin S
\end{align}
for some constant $C$.
